---
format:
    revealjs:
      slide-number: false
      progress: false
      chalkboard: true
      preview-links: true
      theme: [default, custom.scss]
---

```{r setup, include=FALSE}
library(knitr)

opts_chunk$set(echo = FALSE, 
               message = FALSE, 
               warning = FALSE, 
               results = "asis", 
               out.width = "100%")

library(tidyverse)
library(tinytable)
```



::: {style="text-align: center"}
## Inference {.center}

### POLI SCI 210

Introduction to Empirical Methods in Political Science

:::


## AI Prompts {background-color="#B6ACD1"}

TBD

::: aside
**Reminder:** These are just suggestions, feel free to be creative with your prompts!
:::

## Last week

- Theory as explanation

- Elements of a good theory

- Data as the raw material from which we draw inferences

- Types of data (variables)

. . .

::: {.r-stack}
Why do we care so much about data?
:::

## This week: Inference

::: incremental
- We use data to conduct **inference**

- **Inference:** Using what we know to understand something we do not know

- **_Statistical_ inference:** Using the *data* we have to understand something for which we do *not* have data

- Data alone does not help unless we **summarize** it

- **TUESDAY:** *Estimation* and *uncertainty*

- **THURSDAY:** Hypothesis testing

:::

## Running example

:::: {.columns}

::: {.column width="40%"}

```{r}
heights = tribble(
  ~Position, ~`Height (inches)`,
  1, 59.0,
  2, 61.5,
  3, 62.0,
  4, 63.0,
  5, 64.7,
  6, 65.5,
  7, 68.0,
  8, 69.0,
  9, 72.0,
  10, 72.0,
  11, 72.0,
)

heights %>% tt() %>% style_tt(fontsize = 0.7)
```

:::

::: {.column width="60%"}

::: incremental
- How would you describe this group of people?

- Taller/shorter than most?

- How do we summarize this data?
:::

:::


::::

## Running example

:::: {.columns}

::: {.column width="40%"}

```{r}
heights = tribble(
  ~Position, ~`Height (inches)`,
  1, 59.0,
  2, 61.5,
  3, 62.0,
  4, 63.0,
  5, 64.7,
  6, 65.5,
  7, 68.0,
  8, 69.0,
  9, 72.0,
  10, 72.0,
  11, 72.0,
)

heights %>% tt() %>% style_tt(fontsize = 0.7)
```

:::

::: {.column width="60%"}
**Measures of central tendency**

::: incremental
- **Mean:** $\frac{1}{n}\sum_{i=1}^{n} x_i$

- **Median:** Value in the middle

- **Mode:** Value most often repeated
:::

:::

::::